Rapid cycleway prioritisation tool

### 2020-06-05 code/index.Rmd
Fix more typos in index, thanks to kg

---
title: "Rapid cycleway prioritisation tool"
author: "Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds"
output: 
  html_document: 
    toc: yes
bibliography: 
  # - index.bib
  - ../refs.bib
---

@@

These main layers, and two other layers that informed the results but which are not shown be default, are described in detail below.

@@

We used certain eligibility criteria for the Top ranked new cycleways:

- a road which already has an existing cycleway alongside the majority of its length cannot be a top route. 
- unnamed roads cannot be a top route 
- routes less than 500m in length cannot be a top route

Some road sections highlighted by the tool may not currently be suitable for new cycleways.
Road sections that are on strategic and major road networks (SRN and MRN respectively) and which had a high speed limit (40 mph+) before the COVID-19 crises can be in the top routes.
To highlight such roads, the speed limits of road sections are listed in the table of the top ranked new cycleway locations to inform prioritisation and speed reduction interventions.

@@

This layer is represents what a joined-up cycle network could look like if new cycleways were placed along corridors of high potential, by either closing roads to motorised traffic or creating one-way systems.
It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.

The cohesive network layer supports consideration of **long term** plans and alternatives to road space reallocation.
Road sections in the cohesive network but not in the Top ranked new cycleway layer may be suitable candidates for one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
The cohesive network layer was created based on evidence that cohesive cycle networks, that are continuous and do not 'give up' at awkward junctions, are a key component of plans to get people cycling.

@@

- The 'Top ranked new cycleways' highlighted in the map-based results were produced automatically based on a simple score combined cycling potential for school and commute trips.  **The results should be interpreted based on local knowledge and an understanding of the limitations of the data. The results are designed to complement not replace other factors such as feedback from the local community and long-term strategic cycle network plans.**
- The estimates of cycling potential only include school and work travel. Accounting for increases in cycling for other trip purposes such as for shopping and leisure would increase cycling potential substantially.
- The road network data was taken from the Cycling Infrastructure Prioritisation Toolkit (CyIPT) project which was completed in 2017, so may be out of date.
- The results are based on the existing road network and does not flag potential routes that could be constructed, e.g. along old railway tracks or across open areas.
- The results can highlight major roads that have a speed limit above 40mph that may currently be unsuitable for new cycleways created with light segregation. Roads not controlled by local authorities on the SRN can also be highlighted (although motorways are excluded); we do not alter the results based on who is in control of the roads.
- Other than the Cohesive network layer, which highlights what a joined-up cycle network could look like if roads could be made one way to motor traffic and opened-up to cycling via modal filters, the results do not provide evidence on the full range of interventions. Identifying locations for filtered permeability interventions in residential areas, for example, requires data on residential networks not contained in this analysis, for example.

>>>>

---
title: "Rapid cycleway prioritisation tool"
author: "Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds"
output: 
  html_document: 
    toc: yes
bibliography: 
  - index.bib
  - ../refs.bib
---

@@

These main layers, and two other layers that informed the results but which are not shown by default, are described in detail below.

@@

We used the following criteria for the Top ranked new cycleways:

- A road which already has an existing cycleway alongside the majority of its length cannot be a top route.
- Unnamed roads cannot be a top route.
- Routes less than 500m in length cannot be a top route.

Some road sections highlighted by the tool may not currently be suitable for new cycleways.
Road sections that are on strategic and major road networks (SRN and MRN respectively) and which had a high speed limit (40 mph+) before the COVID-19 crisis can be in the top routes.
To highlight such roads, the speed limits of road sections are listed in the table of the top ranked new cycleway locations to inform prioritisation and speed reduction interventions.

@@

This layer represents what a joined-up cycle network could look like if new cycleways were placed along corridors of high potential, by either closing roads to motorised traffic or creating one-way systems.
It shows what a joined-up cycle network could look like if there were no constraints on road space, along some of the most important corridors.

The cohesive network layer supports consideration of **long term** plans and alternatives to road space reallocation.
Road sections in the cohesive network but not in the Top ranked new cycleway layer may be suitable candidates for one way streets, where there may be insufficient space for road space reallocation without changing the navigable network for motor traffic.
The cohesive network layer was created to highlight the need for cohesive cycle network that and do not 'give up' at awkward junctions.
Cohesive networks are a key to getting people cycling [@buehler_bikeway_2016].

@@

- The 'Top ranked new cycleways' highlighted in the map-based results were produced automatically based on a simple score combined cycling potential for school and commute trips.  **The results should be interpreted based on local knowledge and an understanding of the limitations of the data. The results are designed to complement not replace other factors such as feedback from the local community and long-term strategic cycle network plans.**
- The estimates of cycling potential only include school and work travel. Accounting for increases in cycling for other trip purposes such as for shopping and leisure would increase cycling potential substantially.
- The road network data was taken from the Cycling Infrastructure Prioritisation Toolkit (CyIPT) project which was completed in 2017, so may be out of date.
- The results are based on the existing road network. They do not flag potential routes that could be constructed, e.g. along old railway tracks or across open areas.
- The results can highlight major roads that have a speed limit above 40mph that may currently be unsuitable for new cycleways created with light segregation. Roads not controlled by local authorities on the SRN can also be highlighted (although motorways are excluded); we do not alter the results based on who is in control of the roads.
- Other than the Cohesive network layer, which highlights what a joined-up cycle network could look like if roads could be made one way to motor traffic and opened-up to cycling via modal filters, the results do not provide evidence on the full range of interventions. Identifying locations for filtered permeability interventions in residential areas, for example, requires data on residential networks not contained in this analysis.
